---
title: "Assignment 4 - Time Series"
author: "Attolini Maria Carmela, Magi Alessandro"
date: 
header-includes:
output: pdf_document


---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	cache = TRUE,
	include = TRUE
	
)
```

```{r Libraries, echo = FALSE}
library(readxl)
library(janitor)
library(stringr)
library(knitr)
library(tinytex)
library(depmixS4)
library(xts)
library(zoo)
library(psych)
library(lubridate)
library(dlm)
library(tidyverse)
```


# Exercise 2

The dataset Nile, available in R, provides the measurements of the annual flow of the river Nile at Ashwan,
for the period 1871-1970.

## Point 1 

```{r Nile Dataset and Plot, echo = FALSE, eval = TRUE, out.width='70%', fig.show='hold', fig.align='center', fig.cap='Level of the Nile Over Time'}

niledf <- data.frame(nile=as.matrix(Nile), year=time(Nile))

nile_graph <- niledf %>%
  ggplot() + 
  aes(y=nile,
      x=year) + 
  labs(x = "Time",
       y = "Level of Nile") +
  geom_line() +
  theme_minimal()

nile_graph

```


```{r Estimating Model, echo = FALSE, eval = TRUE}

buildrw <- function(parameters){
    dlmModPoly(order=1, dV=parameters[1], dW=parameters[2], m0=Nile[1])}

outMLE <- dlmMLE(Nile, rep(100, 2), buildrw, lower=c(0.000001, 0), hessian=TRUE)   

# Estimating Standard Errors
AsymCov <- solve(outMLE$hessian) 
MLEse <- sqrt(diag(AsymCov))  

rw <- buildrw(outMLE$par)
outFilt <- dlmFilter(Nile,rw)

kable(data.frame(c(outMLE$par[1],MLEse[1]),
                c(outMLE$par[2],MLEse[2]),
                row.names = c("$Value$", "$se$")), 
      digits = 2, 
      caption = "MLE Optimal Estimators and SE", 
      col.names = c("$V$","$W$"), 
      align = 'c')



```

\newpage

##  Point 2

After having estimated the variances for the measurement error $\sigma^2_v$ and the variance for the evolution error $\sigma^2_w$ through MLEs (Table 1), we plug them in the model as if they were "known" parameters. 

\begin{eqnarray*} 
Y_t &= \theta_t + v_t \quad    & v_t \overset{i.i.d.}\sim N(0, 15098.70)\\
\theta_t &= \theta_{t-1} + w_t \quad   & v_t \overset{i.i.d.}\sim N(0,1469.02)
\end{eqnarray*}

```{r Fitted and Std. Dev., echo = FALSE, eval = TRUE, out.width='60%', fig.show='hold', fig.cap='Graphs Point 1', fig.align='center'}

#Adding fitted to dataframe
niledf <- niledf %>%
  mutate(fitted = dropFirst(outFilt$m))

#This is the standard deviation
listC <- dlmSvd2var(outFilt$U.C, outFilt$D.C)
sqrtC <- sqrt(unlist(listC))

#Adding standard dev to dataframe
niledf <- niledf %>%
  mutate(stdev = dropFirst(sqrtC))

#Adding credible interval
niledf <- niledf %>%
  mutate(lower = fitted + qnorm(0.10, sd=stdev),
         upper = fitted + qnorm(0.90, sd=stdev))


graph_fitted <- niledf %>%
  ggplot() + 
  aes( x = year,
       y = fitted) + 
  labs(title= "A: Filtering States Estimates",
       x= "Time") +
  ylab("") +
  geom_point() +
  geom_line(show.legend = FALSE) +
  theme_minimal() +
  geom_ribbon(aes(ymin=niledf$lower,
                  ymax=niledf$upper),
              linetype=2, 
              alpha=0.1)

graph_fitted


```

## Point 3


```{r Graphs, echo = FALSE, eval = TRUE, out.width='60%', fig.show='hold', fig.cap='Graphs Point 3', fig.align='center'}

qqnorm(residuals(outFilt, sd=FALSE))
qqline(residuals(outFilt, sd=FALSE))
tsdiag(outFilt)

```

We are dealing with a univariate time series, hence the sequence of standardized innovation $\Tilde{e_1}$,...,$\Tilde{e_t}$, if the model is correct, should be a Gaussian white noise. Several statistical tests can be carried out on the standardized forecast errors to check if the model is correct. Specifically, they are divided into: 
\begin{enumerate}
    \item [1.] tests to check if $\Tilde{e_t}$'s distribution is a Gaussian standard normal
    \item[2.] tests to check if the standardized innovation are correlated or not.
\end{enumerate}
In an easier way, however, we can plot the standardized forecasts errors and implement a visual analysis.
