---
title: "Final Assignment - Time Series"
author: "Attolini Maria Carmela, Magi Alessandro"
date: 
output: pdf_document
geometry: margin = 0.3in
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	cache = TRUE,
	include = TRUE
)
```

```{r Libraries, echo = FALSE}
library(tidyverse)
library(readxl)
library(janitor)
library(stringr)
library(knitr)
library(tinytex)
library(depmixS4)
library(xts)
library(zoo)
library(psych)
library(lubridate)
```

We are working with space-temporal data from the U.S. Environmental Protection Agency (EPS). The dataset displays hourly air quality data measured in 10 different stations in California from June to September 2020. In particular, we have decided to focus and analyze the data of station 55 (station_id 55). 


```{r Importing Data, echo=FALSE}
path <- "C:\\Users\\magia\\OneDrive - Università Commerciale Luigi Bocconi\\AA Second Semester\\Time Series\\Time Series Magi-Attolini\\Final\\ts_epa_2020_west_sept_fill.csv"
dataset <- read_csv(path, col_types = cols(temp = col_double(), wind = col_double()))

#Here we exctract only id_95 station
id_95 <- dataset %>%
  filter(station_id==55)

```

# Plots and Descriptive Stats

```{r Work the Dataset, echo=FALSE}


#Here I create long dataset to wrap plot
id_95_long <- id_95 %>%
  dplyr::select(datetime, pm25, temp, wind) %>%
  pivot_longer(cols=c("pm25", "temp", "wind"), names_to = "type")

#Here I collapse by day
id_95_coll <- id_95 %>%
  mutate(date=as.Date(datetime), time=format(datetime, "%H:%M:%S")) %>%
  group_by(date) %>%
  summarize(station_id=mean(station_id), pm25=mean(pm25), temp=mean(temp), wind=mean(wind))

#Same for collapsed
id_95_coll_long <- id_95_coll %>%
  dplyr::select(date, pm25, temp, wind) %>%
  pivot_longer(cols=c("pm25", "temp", "wind"), names_to = "type")

#Half of the day
id_95_half <- id_95 %>%
  mutate(date=as.Date(datetime), 
         first_half= ifelse(hour(datetime) < 12 & hour(datetime) >=0, 1, 0)) %>%
  group_by(date, first_half) %>%
  summarize(station_id=mean(station_id), pm25=mean(pm25), temp=mean(temp), wind=mean(wind), .groups = "drop")

id_95_half <- id_95_half %>%
  mutate(time=ifelse(first_half==1, "11:59:99","23:59:99"),
         datetime=as.POSIXct(paste(date, time), "%Y-%m-%d %H:%M:%S"))

#Same for collapsed
id_95_half_long <- id_95_half %>%
  dplyr::select(datetime, pm25, temp, wind) %>%
  pivot_longer(cols=c("pm25", "temp", "wind"), names_to = "type")
      

```


```{r Descriptive Stats, echo = FALSE, eval = TRUE}

#Little check to verify 0 missing values
missing_pm <- id_95 %>%
  filter(is.na(pm25)) %>%
  count()

missing_temp <- id_95 %>%
  filter(is.na(temp)) %>%
  count()

missing_wind <- id_95 %>%
  filter(is.na(wind)) %>%
  count()

#Constructiv Descriptive Statistics Table
descriptive_stats <- id_95 %>%
  dplyr::select(pm25, temp, wind)

descriptive_stats <- describe(descriptive_stats, fast=TRUE, ranges = TRUE, na.rm=TRUE)
descriptive_stats <- descriptive_stats[, 2:7]
row.names(descriptive_stats) <- c("PM2.5", "Temperature", "Wind")

kable(descriptive_stats, 
      digits = 2, 
      caption = "Descriptive Statistics (Hourly Data)", 
      col.names = c("Number of Obs", "Mean", "Standard Deviation", "Min", "Max", "Range"), 
      align = 'c')

#How many times pm25 pass the threshold?

id_95_freq <- id_95 %>%
  dplyr::select(pm25) %>%
  mutate(pass= ifelse(pm25 >= 25, 1, 0)) %>%
  dplyr::summarize(n_obs=n(), n_pass= sum(pass), freq = round(n_pass/n_obs*100,2))

id_95_freq[,3] <- paste0(id_95_freq[1,3],"%")

#id_95_freq <- id_95 %>%
#  dplyr::select(pm25) %>%
#  filter(pm25 >= 25) %>%
#  dplyr::summarize(n_obs= n(), freq = round(n_obs/nrow(id_95)*100,2))

kable(id_95_freq, 
      digits = 2, 
      caption = "PM2.5 Frequency of Passing Threshold (Hourly Data)", 
      col.names = c("N. Obs", "N. Pass", "Frequency"), 
      align = 'c')

```


```{r Hourly Graph, echo=FALSE, eval = TRUE, out.width='50%', fig.show='hold', fig.cap='Time Series by Hour at Station 55'}

id_95_graph <- id_95 %>%
  ggplot() +
  aes(x=datetime,
      y=pm25) +
  labs(title= "A: PM2.5 Level Over Time",
       x= "",
       y= "") + 
  geom_line() + 
  annotate(geom="rect", 
           xmax=max(id_95$datetime), 
           ymin=25,
           xmin=min(id_95$datetime),
           ymax=max(id_95$pm25), 
           fill="darkred",
           alpha=0.2) +
  annotate(geom="text", 
           x=as.POSIXct("2020-06-25 23:00:00 UTC"), 
           y=max(id_95$pm25)-25, 
           label="Dangerous PM2.5 level", 
           color="darkred") +
  theme_minimal()

labels_facet <- c(
                    `pm25` = "PM2.5",
                    `temp` = "Temperature",
                    `wind` = "Wind"
                    )

graph_id_95_long <- id_95_long %>%
  ggplot() + 
  aes( x = datetime,
       y = value,
       color = type) + 
  labs(title= "B: Comparison Between PM25, Temperature and Wind Over Time",
       x= "") +
  ylab("") +
  geom_line(show.legend = FALSE) +
  facet_wrap(~ type, ncol = 1, scale="free_y", labeller = as_labeller(labels_facet))+ 
  theme_minimal()

id_95_graph
graph_id_95_long

```

```{r Half Day Graph, echo=FALSE, eval = TRUE, out.width='50%', fig.show='hold', fig.cap='Time Series Mean Over Half-Day at Station 55'}

id_95_half_graph <- id_95_half %>%
  ggplot() +
  aes(x=datetime,
      y=pm25) +
  labs(title= "A: PM2.5 Level Over Time",
       x= "",
       y= "") + 
  geom_line() + 
  annotate(geom="rect", 
           xmax=max(id_95_half$datetime), 
           ymin=25,
           xmin=min(id_95_half$datetime),
           ymax=max(id_95_half$pm25), 
           fill="darkred",
           alpha=0.2) +
  annotate(geom="text", 
           x=as.POSIXct("2020-06-25 23:00:00 UTC"), 
           y=max(id_95_half$pm25)-25, 
           label="Dangerous PM2.5 level", 
           color="darkred") +
  theme_minimal()



graph_id_95_half_long <- id_95_half_long %>%
  ggplot() + 
  aes( x = datetime,
       y = value,
       color = type) + 
  labs(title= "B: Comparison Between PM25, Temperature and Wind Over Time",
       x= "") +
  ylab("") +
  geom_line(show.legend = FALSE) +
  facet_wrap(~ type, ncol = 1, scale="free_y", labeller = as_labeller(labels_facet))+ 
  theme_minimal()

id_95_half_graph
graph_id_95_half_long

```

```{r Collapsed Graphs, echo=FALSE, eval = TRUE, out.width='50%', fig.show='hold', fig.cap='Time Series Mean Over Day at Station 55'}


id_95_coll_graph <- id_95_coll %>%
  ggplot() +
  aes(x=date,
      y=pm25) +
  labs(title= "A: PM2.5 Level Over Time",
       x= "",
       y= "") + 
  geom_line() + 
  annotate(geom="rect", 
           xmax=max(id_95_coll$date), 
           ymin=25,xmin=min(id_95_coll$date),
           ymax=max(id_95_coll$pm25), 
           fill="darkred",
           alpha=0.2) +
  annotate(geom="text", 
           x=as.Date("2020-06-25"), 
           y=max(id_95_coll$pm25)-10, 
           label="Dangerous PM2.5 level", 
           color="darkred") +
  theme_minimal()

graph_id_95_coll_long <- id_95_coll_long %>%
  ggplot() + 
  aes( x = date,
       y = value,
       color = type) + 
  labs(title= "B: Comparison Between PM25, Temperature and Wind Over Time",
       x= "") +
  ylab("") +
  geom_line(show.legend = FALSE) +
  facet_wrap(~ type, ncol = 1, scale="free_y", labeller = as_labeller(labels_facet))+ 
  theme_minimal()

id_95_coll_graph
graph_id_95_coll_long

```

\newpage



# Model Specification

The data plotted, especially when looking at the average of the observations over a single day, seem to suggest a \textit{non-stationary time series}. Given this, it seems plausible to proceed with the specification of a \textit{Gaussian Hidden Markov Model (HMM)}.

We proceed with the specification of the following model with 4 latent states: 

\begin{equation*}
\begin{cases}
Y_t = \mu_1 + \epsilon_t, \quad \epsilon_t \overset{iid}{\sim} N(0, \sigma_1^2)  
& \text{if the state $S_t=1$} \\
Y_t = \mu_2 + \epsilon_t, \quad \epsilon_t \overset{iid}{\sim} N(0, \sigma_2^2) 
&\text{if the state $S_t=2$} \\
Y_t = \mu_3 + \epsilon_t, \quad \epsilon_t \overset{iid}{\sim} N(0, \sigma_3^2)  
& \text{if the state $S_t=3$} \\
Y_t = \mu_4 + \epsilon_t, \quad \epsilon_t \overset{iid}{\sim} N(0, \sigma_4^2) 
&\text{if the state $S_t=4$}.
\end{cases}
\end{equation*}

\newpage


# Optimal Estimate: Daily Data


```{r HMM estimation Day, echo=FALSE, include=FALSE}
set.seed(2022)
model <- depmix(id_95_coll$pm25 ~ 1, data=id_95_coll, nstates=4)
fmodel <- fit(model)
summary(fmodel)

```

```{r MLE Estimation Day, echo=FALSE}

MLEse <- standardError(fmodel) %>%
  mutate(par = round(par,4),
         se = round(se, 4))

```


```{r Initial Prob Day, echo=FALSE}

in_prob <- data.frame(0,0,0,0)
in_prob[nrow(in_prob),] <- MLEse[1:4,1]
in_prob[nrow(in_prob)+1,] <- MLEse[1:4,3]

in_prob <- data.frame(in_prob, row.names = c("Probability", "$se$"))

kable(in_prob, 
      digits = 4, 
      caption = "Initial Probability",
      col.names = c("Prob. 1", "Prob. 2", "Prob. 3", "Prob. 4"),
      align = 'c')

```

```{r Transition Matrix Day, echo=FALSE}

trans_mat <- data.frame(0,0,0,0)
trans_mat[nrow(trans_mat),] <- MLEse[5:8,1]
trans_mat[nrow(trans_mat) + 1,] <- MLEse[5:8,3]
trans_mat[nrow(trans_mat) + 1,] <- MLEse[9:12,1]
trans_mat[nrow(trans_mat) + 1,] <- MLEse[9:12,3]
trans_mat[nrow(trans_mat) + 1,] <- MLEse[13:16,1]
trans_mat[nrow(trans_mat) + 1,] <- MLEse[13:16,3]
trans_mat[nrow(trans_mat) + 1,] <- MLEse[17:20,1]
trans_mat[nrow(trans_mat) + 1,] <- MLEse[17:20,3]

trans_mat <- data.frame(trans_mat, row.names = c("From State 1", "$se_{1}$","From State 2", "$se_{2}$","From State 3", "$se_{3}$","From State 4", "$se_{4}$"))

trans_mat[2,] <- paste0("(", format(unlist(trans_mat[2,])),")")
trans_mat[4,] <- paste0("(", format(unlist(trans_mat[4,])),")")
trans_mat[6,] <- paste0("(", format(unlist(trans_mat[6,])),")")
trans_mat[8,] <- paste0("(", format(unlist(trans_mat[8,])),")")

kable(trans_mat, 
      digits = 4, 
      caption = "Transition Matrix",
      col.names = c("To 1", "To 2", "To 3", "To 4"),
      align = 'c')

```

```{r Emission Distribution Day, echo=FALSE}

em_dist <- data.frame(0,0,0,0)
em_dist[nrow(em_dist),] <- MLEse[c(21,23,25,27),1]
em_dist[nrow(em_dist) + 1,] <- MLEse[c(21,23,25,27),3]
em_dist[nrow(em_dist) + 1,] <- MLEse[c(22,24,26,28),1]
em_dist[nrow(em_dist) + 1,] <- MLEse[c(22,24,26,28),3]


em_dist <- data.frame(em_dist, row.names = c("$\\mu$", "$se_{\\mu}$", "$\\sigma$", "$se_\\sigma$"))

em_dist[2,] <- paste0("(", format(unlist(em_dist[2,])),")")
em_dist[4,] <- paste0("(", format(unlist(em_dist[4,])),")")

kable(em_dist, 
      digits = 4, 
      caption = "Guassian Emission Distribution",
      col.names = c("State 1", "State 2", "State 3", "State 4"),
      align = 'c')

```

```{r Graph 1 Day, echo = FALSE, eval = TRUE, out.width='50%',fig.align='center', fig.cap='Add Caption'}

estStates <- posterior(fmodel)

id_95_states <- id_95_coll %>%
  mutate(states=estStates[,1])

id_95_states_long <- id_95_states %>%
  pivot_longer(cols=c("pm25", "states"), names_to = "type")

graph_states <- id_95_states_long %>%
  ggplot() + 
  aes( x = date,
       y = value,
       color = type) + 
  labs(x= "", title = "B: PM2.5 Level and Optimal State Process") +
  ylab("") +
  geom_line(show.legend = FALSE) +
  facet_wrap(~ type, ncol = 1, scale="free_y") + 
  theme_minimal()

```

```{r Graph 2 Day, echo = FALSE, eval = TRUE}


estMean1=fmodel@response[[1]][[1]]@parameters$coefficients
estMean2=fmodel@response[[2]][[1]]@parameters$coefficients
estMean3=fmodel@response[[3]][[1]]@parameters$coefficients
estMean4=fmodel@response[[4]][[1]]@parameters$coefficients
estMeans=rep(estMean1, nrow(id_95_coll))
estMeans[estStates[,1]==2]=estMean2
estMeans[estStates[,1]==3]=estMean3
estMeans[estStates[,1]==4]=estMean4

id_95_states_means <- id_95_states %>%
  mutate(means= estMeans)

id_95_states_means_long <- id_95_states_means %>%
  pivot_longer(cols=c("pm25", "means"), names_to = "type")

graph_means <- id_95_states_means_long %>%
  ggplot() + 
  aes( x = date,
       y = value,
       color = type) + 
  labs(x= "", title = "A: PM2.5 Level with Period Specific Means") +
  ylab("") +
  geom_line(show.legend = FALSE) + 
  theme_minimal()


```

```{r Graphs HMM Day, echo = FALSE, eval = TRUE, out.width='50%', fig.show='hold', fig.cap='HMM Visualization Daily Data'}

graph_means
graph_states

```


# Appendix

## A1 HMM Half Day

```{r HMM estimation Half, echo=FALSE, include=FALSE}
set.seed(2022) #With this seed the higher state is 3, the lower is 1 and the middle is 2
model <- depmix(id_95_half$pm25 ~ 1, data=id_95_half, nstates=4)
fmodel <- fit(model)
summary(fmodel)

```

```{r MLE Estimation Half, echo=FALSE}

MLEse <- standardError(fmodel) %>%
  mutate(par = round(par,4),
         se = round(se, 4))

```

```{r Initial Prob Half, echo=FALSE}

in_prob <- data.frame(0,0,0,0)
in_prob[nrow(in_prob),] <- MLEse[1:4,1]
in_prob[nrow(in_prob)+1,] <- MLEse[1:4,3]

in_prob <- data.frame(in_prob, row.names = c("Probability", "$se$"))

kable(in_prob, 
      digits = 4, 
      caption = "Initial Probability",
      col.names = c("Prob. 1", "Prob. 2", "Prob. 3", "Prob. 4"),
      align = 'c')

```

```{r Transition Matrix Half, echo=FALSE}

trans_mat <- data.frame(0,0,0,0)
trans_mat[nrow(trans_mat),] <- MLEse[5:8,1]
trans_mat[nrow(trans_mat) + 1,] <- MLEse[5:8,3]
trans_mat[nrow(trans_mat) + 1,] <- MLEse[9:12,1]
trans_mat[nrow(trans_mat) + 1,] <- MLEse[9:12,3]
trans_mat[nrow(trans_mat) + 1,] <- MLEse[13:16,1]
trans_mat[nrow(trans_mat) + 1,] <- MLEse[13:16,3]
trans_mat[nrow(trans_mat) + 1,] <- MLEse[17:20,1]
trans_mat[nrow(trans_mat) + 1,] <- MLEse[17:20,3]

trans_mat <- data.frame(trans_mat, row.names = c("From State 1", "$se_{1}$","From State 2", "$se_{2}$","From State 3", "$se_{3}$","From State 4", "$se_{4}$"))

trans_mat[2,] <- paste0("(", format(unlist(trans_mat[2,])),")")
trans_mat[4,] <- paste0("(", format(unlist(trans_mat[4,])),")")
trans_mat[6,] <- paste0("(", format(unlist(trans_mat[6,])),")")
trans_mat[8,] <- paste0("(", format(unlist(trans_mat[8,])),")")

kable(trans_mat, 
      digits = 4, 
      caption = "Transition Matrix",
      col.names = c("To 1", "To 2", "To 3", "To 4"),
      align = 'c')

```

```{r Emission Distribution Half, echo=FALSE}

em_dist <- data.frame(0,0,0,0)
em_dist[nrow(em_dist),] <- MLEse[c(21,23,25,27),1]
em_dist[nrow(em_dist) + 1,] <- MLEse[c(21,23,25,27),3]
em_dist[nrow(em_dist) + 1,] <- MLEse[c(22,24,26,28),1]
em_dist[nrow(em_dist) + 1,] <- MLEse[c(22,24,26,28),3]


em_dist <- data.frame(em_dist, row.names = c("$\\mu$", "$se_{\\mu}$", "$\\sigma$", "$se_\\sigma$"))

em_dist[2,] <- paste0("(", format(unlist(em_dist[2,])),")")
em_dist[4,] <- paste0("(", format(unlist(em_dist[4,])),")")

kable(em_dist, 
      digits = 4, 
      caption = "Guassian Emission Distribution",
      col.names = c("State 1", "State 2", "State 3", "State 4"),
      align = 'c')

```

```{r Graph 1 Half, echo = FALSE, eval = TRUE, out.width='50%',fig.align='center', fig.cap='Add Caption'}

estStates <- posterior(fmodel)

id_95_states <- id_95_half %>%
  mutate(states=estStates[,1])

id_95_states_long <- id_95_states %>%
  pivot_longer(cols=c("pm25", "states"), names_to = "type")

graph_states <- id_95_states_long %>%
  ggplot() + 
  aes( x = datetime,
       y = value,
       color = type) + 
  labs(x= "", title = "B: PM2.5 Level and Optimal State Process") +
  ylab("") +
  geom_line(show.legend = FALSE) +
  facet_wrap(~ type, ncol = 1, scale="free_y") + 
  theme_minimal()

```

```{r Graph 2 Half, echo = FALSE, eval = TRUE}


estMean1=fmodel@response[[1]][[1]]@parameters$coefficients
estMean2=fmodel@response[[2]][[1]]@parameters$coefficients
estMean3=fmodel@response[[3]][[1]]@parameters$coefficients
estMean4=fmodel@response[[4]][[1]]@parameters$coefficients
estMeans=rep(estMean1, nrow(id_95_half))
estMeans[estStates[,1]==2]=estMean2
estMeans[estStates[,1]==3]=estMean3
estMeans[estStates[,1]==4]=estMean4

id_95_states_means <- id_95_states %>%
  mutate(means= estMeans)

id_95_states_means_long <- id_95_states_means %>%
  pivot_longer(cols=c("pm25", "means"), names_to = "type")

graph_means <- id_95_states_means_long %>%
  ggplot() + 
  aes( x = datetime,
       y = value,
       color = type) + 
  labs(x= "", title = "A: PM2.5 Level with Period Specific Means") +
  ylab("") +
  geom_line(show.legend = FALSE) + 
  theme_minimal()


```

```{r Graphs HMM Half, echo = FALSE, eval = TRUE, out.width='50%', fig.show='hold', fig.cap='HMM Visualization Half Day Mean'}

graph_means
graph_states

```

\newpage

## A2 HMM Hourly

```{r HMM estimation Hour, echo=FALSE, include=FALSE}
set.seed(2022) #With this seed the higher state is 3, the lower is 1 and the middle is 2
model <- depmix(id_95$pm25 ~ 1, data=id_95, nstates=4)
fmodel <- fit(model)
summary(fmodel)

```

```{r MLE Estimation Hour, echo=FALSE}

MLEse <- standardError(fmodel) %>%
  mutate(par = round(par,4),
         se = round(se, 4))

```

```{r Initial Prob Hour, echo=FALSE}

in_prob <- data.frame(0,0,0,0)
in_prob[nrow(in_prob),] <- MLEse[1:4,1]
in_prob[nrow(in_prob)+1,] <- MLEse[1:4,3]

in_prob <- data.frame(in_prob, row.names = c("Probability", "$se$"))

kable(in_prob, 
      digits = 4, 
      caption = "Initial Probability",
      col.names = c("Prob. 1", "Prob. 2", "Prob. 3", "Prob. 4"),
      align = 'c')

```

```{r Transition Matrix Hour, echo=FALSE}

trans_mat <- data.frame(0,0,0,0)
trans_mat[nrow(trans_mat),] <- MLEse[5:8,1]
trans_mat[nrow(trans_mat) + 1,] <- MLEse[5:8,3]
trans_mat[nrow(trans_mat) + 1,] <- MLEse[9:12,1]
trans_mat[nrow(trans_mat) + 1,] <- MLEse[9:12,3]
trans_mat[nrow(trans_mat) + 1,] <- MLEse[13:16,1]
trans_mat[nrow(trans_mat) + 1,] <- MLEse[13:16,3]
trans_mat[nrow(trans_mat) + 1,] <- MLEse[17:20,1]
trans_mat[nrow(trans_mat) + 1,] <- MLEse[17:20,3]

trans_mat <- data.frame(trans_mat, row.names = c("From State 1", "$se_{1}$","From State 2", "$se_{2}$","From State 3", "$se_{3}$","From State 4", "$se_{4}$"))

trans_mat[2,] <- paste0("(", format(unlist(trans_mat[2,])),")")
trans_mat[4,] <- paste0("(", format(unlist(trans_mat[4,])),")")
trans_mat[6,] <- paste0("(", format(unlist(trans_mat[6,])),")")
trans_mat[8,] <- paste0("(", format(unlist(trans_mat[8,])),")")

kable(trans_mat, 
      digits = 4, 
      caption = "Transition Matrix",
      col.names = c("To 1", "To 2", "To 3", "To 4"),
      align = 'c')

```

```{r Emission Distribution Hour, echo=FALSE}

em_dist <- data.frame(0,0,0,0)
em_dist[nrow(em_dist),] <- MLEse[c(21,23,25,27),1]
em_dist[nrow(em_dist) + 1,] <- MLEse[c(21,23,25,27),3]
em_dist[nrow(em_dist) + 1,] <- MLEse[c(22,24,26,28),1]
em_dist[nrow(em_dist) + 1,] <- MLEse[c(22,24,26,28),3]


em_dist <- data.frame(em_dist, row.names = c("$\\mu$", "$se_{\\mu}$", "$\\sigma$", "$se_\\sigma$"))

em_dist[2,] <- paste0("(", format(unlist(em_dist[2,])),")")
em_dist[4,] <- paste0("(", format(unlist(em_dist[4,])),")")

kable(em_dist, 
      digits = 4, 
      caption = "Guassian Emission Distribution",
      col.names = c("State 1", "State 2", "State 3", "State 4"),
      align = 'c')

```

```{r Graph 1 Hour, echo = FALSE, eval = TRUE, out.width='50%',fig.align='center', fig.cap='Add Caption'}

estStates <- posterior(fmodel)

id_95_states <- id_95 %>%
  mutate(states=estStates[,1])

id_95_states_long <- id_95_states %>%
  pivot_longer(cols=c("pm25", "states"), names_to = "type")

graph_states <- id_95_states_long %>%
  ggplot() + 
  aes( x = datetime,
       y = value,
       color = type) + 
  labs(x= "", title = "B: PM2.5 Level and Optimal State Process") +
  ylab("") +
  geom_line(show.legend = FALSE) +
  facet_wrap(~ type, ncol = 1, scale="free_y") + 
  theme_minimal()

```

```{r Graph 2 Hour, echo = FALSE, eval = TRUE}


estMean1=fmodel@response[[1]][[1]]@parameters$coefficients
estMean2=fmodel@response[[2]][[1]]@parameters$coefficients
estMean3=fmodel@response[[3]][[1]]@parameters$coefficients
estMean4=fmodel@response[[4]][[1]]@parameters$coefficients
estMeans=rep(estMean1, nrow(id_95))
estMeans[estStates[,1]==2]=estMean2
estMeans[estStates[,1]==3]=estMean3
estMeans[estStates[,1]==4]=estMean4

id_95_states_means <- id_95_states %>%
  mutate(means= estMeans)

id_95_states_means_long <- id_95_states_means %>%
  pivot_longer(cols=c("pm25", "means"), names_to = "type")

graph_means <- id_95_states_means_long %>%
  ggplot() + 
  aes( x = datetime,
       y = value,
       color = type) + 
  labs(x= "", title = "A: PM2.5 Level with Period Specific Means") +
  ylab("") +
  geom_line(show.legend = FALSE) + 
  theme_minimal()


```

```{r Graphs HMM Hour, echo = FALSE, eval = TRUE, out.width='50%', fig.show='hold', fig.cap='HMM Visualization Hourly Data'}

graph_means
graph_states

```
